---
title: "M3P_StatisticalAnalysis"
author: "Elise"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/elise/Documents/Mémoire/Main/Data/Templates/M3P")
```

```{r include=FALSE}
WDIR <- "C:/Users/elise/Documents/Mémoire/Main/Data"
subdir <- 'Extracted'
datadir <- sprintf( '%s/%s', WDIR, subdir)

library(statgenHTP)
library(ggplot2)
library(gridExtra)
library(dplyr)
library(lubridate)
library(readxl)
library(janitor)
library(flextable)
library(skimr)
library(rstatix)
library(visreg)
library(lsmeans)
library(car)
library(scales)
library(viridis)
library(ggforce)
library(asreml)
library(patchwork)
library(lme4)
library(lmerTest)
library(desplot)
# Import functions
source("~/Mémoire/Main/Source/functions.R")


list.files()

plant_info <- read.table("plant_info.txt", header = TRUE, sep = "\t")
endpoint <- read.table("endpoint.txt", header = TRUE, sep = "\t")

# plant_info
plant_info <- lapply(plant_info, factor)

# endpoint
matching_cols <- intersect(names(endpoint), names(plant_info))
endpoint[, matching_cols] <- lapply(endpoint[, matching_cols], factor)
endpoint$Date <- date(endpoint$Date)
endpoint$Timestamp <- NA




variables <- c("DW_plant_g")





endpoint_clean <- endpoint
# Run the function on the dataset for all the variables
endpoint_clean <- detect_replace_outliers_by_genotype(endpoint_clean)


endpoint$Plant_type <- substr(endpoint$Genotype, nchar(as.character(endpoint$Genotype)), nchar(as.character(endpoint$Genotype)))
endpoint_clean$Plant_type <- substr(endpoint_clean$Genotype, nchar(as.character(endpoint_clean$Genotype)), nchar(as.character(endpoint_clean$Genotype)))
```



# Statistical modeling for phenotypic traits

The explanatory variable (X) refers to the genotype, which is a categorical variable. The response variables (Y) are the phenotypic data of the variables dataframe.

In this case, the variables are:

And the genotypes are:

```{r}
print(variables)
unique(endpoint$Genotype)
```

### 1. First linear models
Firstly, we model the Y = X + r + c + e
Where 
- Y is the phenotypic trait;
- X the genotype (fixed effect);
- r the row effect (fixed or random);
- c the column effect (fixed or random);

```{r}
fit_models_fixed <- function(data, trait_name) {
  fixed_formula <- as.formula(paste(trait_name, "~ Genotype + Row + Column"))
  fixed_model <- lm(fixed_formula, data)
  print(paste("Summary for fixed effects model of", trait_name))
  print(summary(fixed_model))
  print(anova(fixed_model))
}

fit_models_random <- function(data, trait_name)  {
  random_formula <- as.formula(paste(trait_name, "~ Genotype + (1|Row) + (1|Column)"))
  random_model <- lmer(random_formula, data)
  print(paste("Summary for random effects model of", trait_name))
  print(summary(random_model))
  print(anova(random_model))
  print(ranova(random_model))
}

for (trait in variables) {
  fit_models_fixed(endpoint_clean, trait)
}

for (trait in variables) {
  fit_models_random(endpoint_clean, trait)
}
```

### 2. Linear models with Plant_type
Model with X as Plant_type instead of Genotype, and row and column effects as random effects.
Plant_type is defined as H for Hybrid, L for pure Line and T for Tester. 

```{r}
endpoint_clean$Plant_type <- as.factor(endpoint_clean$Plant_type)
endpoint_clean$Plant_type <- relevel(endpoint_clean$Plant_type, ref = "T") # T as base level

fit_model_plant_type <- function(data, trait) {
  # Random effects model with Plant_type as a fixed effect
  model_formula <- as.formula(paste(trait, "~ Plant_type + (1|Row) + (1|Column)"))
  model <- lmer(model_formula, data)
  print(paste("Summary for mixed effects model of", trait))
  print(summary(model))
  print(anova(model))
  print(ranova(model))
}

for (trait in variables) {
  fit_model_plant_type(endpoint_clean, trait)
}
```

### 3. Linear models with asreml library
```{r}
modasreml <- asreml(fixed = DW_plant_g ~ Genotype,
                    random = ~ Row + Column,
                    residual = ~  NULL,
                    data = endpoint_clean)

plot(modasreml)

summary(modasreml)$varcomp
```

### 4. Linear models with Soil variable
Model with Soil as explicative variable.

```{r eval=FALSE, include=FALSE}
fit_models_for_soil_genotype <- function(data, trait) {
  # Base model with Soil as fixed effect and Row, Column as random effects
  mod0 <- lmer(as.formula(paste(trait, "~ Soil + (1|Row) + (1|Column)")), data = data)
  print(paste("Summary for base model of", trait))
  print(summary(mod0))
  print("Random effects ANOVA for base model:")
  print(ranova(mod0))
  
  # Model with Genotype and Soil interaction
  mod1 <- lmer(as.formula(paste(trait, "~ Genotype * Soil + (1|Row) + (1|Column)")), data = data)
  print(paste("Summary for interaction model of", trait))
  print(summary(mod1))
  print("Random effects ANOVA for interaction model:")
  print(ranova(mod1))
  
  # Model with Genotype and Soil without interaction
  mod2 <- lmer(as.formula(paste(trait, "~ Genotype + Soil + (1|Row) + (1|Column)")), data = data)
  print(paste("Summary for no interaction model of", trait))
  print(summary(mod2))
  print("Random effects ANOVA for no interaction model:")
  print(ranova(mod2))
}

# Iterate through each variable and fit models
for (trait in variables) {
  fit_models_for_soil_genotype(endpoint_clean, trait)
}
```

### 5. Linear models with Soil variable with Plant_type

```{r}
fit_models_for_soil_planttype <- function(data, trait) {
  # Base model with Soil as fixed effect and Row, Column as random effects
  mod0 <- lmer(as.formula(paste(trait, "~ Soil + (1|Row) + (1|Column)")), data = data)
  print(paste("Summary for base model of", trait))
  print(summary(mod0))
  print("Random effects ANOVA for base model:")
  print(ranova(mod0))
  
  # Model with Plant_type and Soil interaction
  mod1 <- lmer(as.formula(paste(trait, "~ Plant_type * Soil + (1|Row) + (1|Column)")), data = data)
  print(paste("Summary for interaction model of", trait))
  print(summary(mod1))
  print("Random effects ANOVA for interaction model:")
  print(ranova(mod1))
  
  # Model with Plant_type and Soil without interaction
  mod2 <- lmer(as.formula(paste(trait, "~ Plant_type + Soil + (1|Row) + (1|Column)")), data = data)
  print(paste("Summary for no interaction model of", trait))
  print(summary(mod2))
  print("Random effects ANOVA for no interaction model:")
  print(ranova(mod2))
}

# Iterate through each variable and fit models
for (trait in variables) {
  fit_models_for_soil_planttype(endpoint_clean, trait)
}
```

