---
title: "FZJ Data Preparation"
author: "Elise"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/elise/Documents/Mémoire/Main/Data/Drive/FZJ")
```

Set the right working directory.
```{r}
setwd("C:/Users/elise/Documents/Mémoire/Main/Data/Drive/FZJ")
```

# Packages importation 

```{r include=FALSE}
rm(list = ls())

library(statgenHTP)
library(ggplot2)
library(gridExtra)
library(dplyr)
library(lubridate)
library(readxl)
library(janitor)
library(flextable)
library(skimr)
library(rstatix)
library(visreg)
library(lsmeans)
library(car)
library(scales)
library(viridis)
library(ggforce)
library(asreml)
library(patchwork)
library(lme4)
library(lmerTest)
library(desplot)
library(purrr)
# Import functions
source("~/Mémoire/functions.R")
```

# 1. Data importation

The first step in this data preparation process involves importing all the pertinent datasets listed in the Google Sheets "Variables template" document. Fist we find the files, then import them.

```{r echo=FALSE}
# List the files
list.files()
```


```{r Data importation, echo=FALSE}
data <- read.delim("multisensor.txt")
data_environment <- read.delim("env_multisensor.txt")
```

We can extract the coordinates of each plant with the ISA_EPPN.xlsx dataset, using a made-up function "coordinates_isaTAB".

```{r}
# Get the coordinates
isaTAB <- read_excel("ISA_EPPN2020_FZJ.xlsx", sheet = "s_exp")
coordinates <- coordinates_isaTAB(isaTAB)
```

### A. Datasets structures
We can take a quick look at all the datasets.

- coordinates
- data
- data_environment

```{r Quick look at the datasets}
head(coordinates)
head(data)
head(data_environment)
```

### B. Data manipulation

This next step standardizes diverse datasets by renaming variables for consistency, converting data into appropriate units, adding necessary columns, and merging the datasets.


```{r}
################################################################################
# COORDINATES
################################################################################
# Unit.ID 
coordinates$Unit.ID <- seq_len(nrow(coordinates))
# Reference for Sample.Name et Unit.ID
reference <- coordinates[, c("Sample.Name", "Unit.ID")]
## We can then copy dataset2$Unit.ID <- reference$Unit.ID[match(dataset2$Sample.Name, reference$Sample.Name)]
# Genotype
reference$Genotype <- isaTAB$`Source Name`

################################################################################
# DATA
################################################################################
# separate for endpoint and timeseries
FW_shoot_g <- data[data$Variable.ID == "Shoot_fresh_weight", ]
FW_shoot_g$Date <- as.Date("2019-11-11")
FW_shoot_g$unit <- reference$Unit.ID[match(FW_shoot_g$Unit.ID, reference$Sample.Name)]
FW_shoot_g$Unit.ID <- FW_shoot_g$unit

DW_shoot_g <- data[data$Variable.ID == "Shoot_dry_weight", ]
DW_shoot_g$Date <- as.Date("2019-11-11")
DW_shoot_g$unit <- reference$Unit.ID[match(DW_shoot_g$Unit.ID, reference$Sample.Name)]
DW_shoot_g$Unit.ID <- DW_shoot_g$unit

DW_root_g <- data[data$Variable.ID == "Root_dry_weight", ]
DW_root_g$Date <- as.Date("2019-11-11")
DW_root_g$unit <- reference$Unit.ID[match(DW_root_g$Unit.ID, reference$Sample.Name)]
DW_root_g$Unit.ID <- DW_root_g$unit

Root_length_cm <- data[data$Variable.ID == "Total_root_length_scan", ]
Root_length_cm$Date <- as.Date("2019-11-11")
Root_length_cm$unit <- reference$Unit.ID[match(Root_length_cm$Unit.ID, reference$Sample.Name)]
Root_length_cm$Unit.ID <- Root_length_cm$unit

Root_angle <- data[data$Variable.ID == "Root_angle", ]
Root_angle$Date <- as.Date("2019-11-11")
Root_angle$unit <- reference$Unit.ID[match(Root_angle$Unit.ID, reference$Sample.Name)]
Root_angle$Unit.ID <- Root_angle$unit

Root_number <- data[data$Variable.ID == "Root_number", ]
Root_number$Date <- as.Date("2019-11-11")
Root_number$unit <- reference$Unit.ID[match(Root_number$Unit.ID, reference$Sample.Name)]
Root_number$Unit.ID <- Root_number$unit


Leaf_number <- data[data$Variable.ID == "Leaf_number", ]
Leaf_number$Date <- gsub("_", "-", Leaf_number$Timestamp)
Leaf_number$Date <- as.Date(Leaf_number$Date)
Leaf_number$unit <- reference$Unit.ID[match(Leaf_number$Unit.ID, reference$Sample.Name)]
Leaf_number$Unit.ID <- Leaf_number$unit

Plant_height_cm <- data[data$Variable.ID == "Plant_height", ]
Plant_height_cm$Date <- gsub("_", "-", Plant_height_cm$Timestamp)
Plant_height_cm$Date <- as.Date(Plant_height_cm$Date)
Plant_height_cm$unit <- reference$Unit.ID[match(Plant_height_cm$Unit.ID, reference$Sample.Name)]
Plant_height_cm$Unit.ID <- Plant_height_cm$unit
```


# 2. Data template

### A. Data template: plant_info
This dataset contains information about the plant: Unit.ID, genotype, replication, row and column location in the greenhouse, and soil treatment.

```{r Template plant_info, include=FALSE}
plant_info <- data.frame(Unit.ID = factor(coordinates$Unit.ID),
                 Genotype = factor(reference$Genotype),
                 Soil = NA,
                 Replication = factor(coordinates$rep),
                 Row = factor(coordinates$nrow),
                 Column = factor(coordinates$ncol),
                 Platform = "FZJ"
                 )
```

### B. Data template: endpoint
This datasets contains information of the end of the experiment (variables at harvest).
It is then linked by the Unit.ID to the plant_info data template.

```{r Template endpoint, include=FALSE}
endpoint <- data.frame(Unit.ID = factor(DW_root_g$Unit.ID),
                       Time = NA, 
                       Date = date(DW_root_g$Date),
                       Timestamp = NA,
                       
                       DW_shoot_g = as.numeric(DW_shoot_g$Value),
                       FW_shoot_g = as.numeric(FW_shoot_g$Value),
                       DW_root_g = as.numeric(DW_root_g$Value),
                       FW_root_g = NA,
                       Leaf_number= NA,
                       Plant_height_cm = NA,
                       DW_plant_g = NA,
                       Root_length_cm = as.numeric(Root_length_cm$Value),
                       Root_number = as.numeric(Root_number$Value),
                       Root_angle = as.numeric(Root_angle$Value),
                       Total_wu = NA,
                       DW_seed_g = NA,
                       FW_seed_g = NA,
                       Leaf_area_cmsquared = NA
)
endpoint <- left_join(endpoint, plant_info, by = "Unit.ID")
```

### C. Data template: timeseries
This section in divided in three data templates: 

- timeseries

- S_timeseries (variables computed from sideview imaging or image processing)

- T_timeseries (variables computed from topview imaging or image processing)

The time interval between data timestamps varies in each platform.
They are then linked by the Unit.ID to the plant_info data template. 


```{r Template timeseries S_timeseries T_timeseries, include=FALSE}

timeseries <- data.frame(Unit.ID = factor(Plant_height_cm$Unit.ID),
                         Time = NA,
                         Date = date(Plant_height_cm$Date),
                         Timestamp = NA,
                         
                         Manual_Plant_height_cm = as.numeric(Plant_height_cm$Value), 
                         Leaf_number = as.numeric(Leaf_number$Value),
                         Wue = NA,
                         Plant_biomass = NA,
                         Ligulated_leaf_number = NA,
                         Plant_emergence = NA,
                         Plant_transpiration = NA,
                         Daily_wu = NA,
                         Soil_water_potential = NA
)
timeseries <- left_join(timeseries, plant_info, by = "Unit.ID")
                      

S_timeseries <- data.frame(Unit.ID = NA,
                           Timestamp = NA,
                           Date = NA,
                           Time = NA,
                           
                           S_Height_cm = NA,
                           S_Height_pixel = NA,
                           S_Area_cmsquared = NA,
                           S_Area_pixel = NA,
                           S_Perimeter_cm = NA,
                           S_Perimeter_pixel = NA,
                           S_Convex_hull_area_cmsquared = NA,
                           S_Solidity = NA,
                           S_Compactness = NA,
                           S_Width_cm = NA,
                           S_Width_pixel = NA,
                           S_Leaf_area_cmsquared = NA
)
S_timeseries <- left_join(S_timeseries, plant_info, by = "Unit.ID")


T_timeseries <- data.frame(Unit.ID = NA,
                           Time = NA, 
                           Date = NA,
                           Timestamp = NA,
                           T_Area_cm_squared = NA,
                           T_Area_pixel = NA,
                           T_Perimeter_cm = NA,
                           T_Perimeter_pixel = NA,
                           T_Convex_hull_area_cmsquared = NA,
                           T_Solidity = NA,
                           T_Compactness = NA,
                           T_Roundness = NA,
                           T_Roundness2 = NA,
                           T_Isotropy = NA,
                           T_Eccentricity = NA,
                           T_Rms = NA,
                           T_Sol = NA
)
T_timeseries <- left_join(T_timeseries, plant_info, by = "Unit.ID")

```

### D. FZJ data templates

- plant_info
- endpoint
- timeseries
- S_timeseries
- T_timeseries

```{r Data template, echo=FALSE}
head(plant_info)
head(endpoint)
head(timeseries)
head(S_timeseries)
head(T_timeseries)
```

# 3. Export the data templates in .txt
Stock the new data sets in a new folder.

```{r Export}
setwd("C:/Users/elise/Documents/Mémoire/Main/Data/Templates/FZJ")

write.table(plant_info, file = "plant_info.txt", sep = "\t", row.names = FALSE, quote = FALSE)
write.table(endpoint, file = "endpoint.txt", sep = "\t", row.names = FALSE, quote = FALSE)
write.table(timeseries, file = "timeseries.txt", sep = "\t", row.names = FALSE, quote = FALSE)
write.table(S_timeseries, file = "S_timeseries.txt", sep = "\t", row.names = FALSE, quote = FALSE)
write.table(T_timeseries, file = "T_timeseries.txt", sep = "\t", row.names = FALSE, quote = FALSE)
```








