---
title: "UCPH Data Preparation"
author: "Elise"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/elise/Documents/Mémoire/Main/Data/Drive/UCPH")
```

Set the right working directory.
```{r}
setwd("C:/Users/elise/Documents/Mémoire/Main/Data/Drive/UCPH")
```

# Packages importation 

```{r include=FALSE}
rm(list = ls())

library(statgenHTP)
library(ggplot2)
library(gridExtra)
library(dplyr)
library(lubridate)
library(readxl)
library(janitor)
library(flextable)
library(skimr)
library(rstatix)
library(visreg)
library(lsmeans)
library(car)
library(scales)
library(viridis)
library(ggforce)
library(asreml)
library(patchwork)
library(lme4)
library(lmerTest)
library(desplot)
library(purrr)
# Import functions
source("~/Mémoire/functions.R")
```

# 1. Data importation

The first step in this data preparation process involves importing all the pertinent datasets listed in the Google Sheets "Variables template" document. Fist we find the files, then import them.

```{r Data files, echo=FALSE}
# List the files
list.files()
```


```{r Data importation, echo=FALSE}

data_pheno <- read_excel("EPPN_Biomass_PhenoLab.xlsx", skip = 1)
data_pheno <- data_pheno[data_pheno$Fixture_number <= 108, ] # The 109etc are not part of the experiment

data_imaging <- read_excel("Cross Platform_Automatic_Height.xlsx")
data_imaging <- data_imaging[data_imaging$`Fixture ID:` <= 108, ] # The 109etc are not part of the experiment

#data_environment <- read.delim("2020_NaPPI_env_multisensor.txt")
```

We can extract the coordinates of each plant with the ISA_EPPN.xlsx dataset, using a made-up function "coordinates_isaTAB".
For UCPH, the coordinates are not in the ISA_EPPN dataset. 

```{r}
design <- read.csv("DesignDIGGER_PhenoLab_JointPf.csv")
```

### A. Datasets structures
We can take a quick look at all the datasets.

- design
- data_pheno
- data_imaging

```{r Quick look at the datasets}
head(design)
head(data_pheno)
head(data_imaging)
#head(data_environment)
```

### B. Data manipulation

This next step standardizes diverse datasets by renaming variables for consistency, converting data into appropriate units, adding necessary columns, and merging the datasets.


```{r}
################################################################################
# DESIGN
################################################################################
# Unit.ID 
design <- rename(design,
                 Unit.ID = UNIT,
                 nrow = ROW,
                 ncol = RANGE,
                 rep = REP,
                 Soil = Substrate)

################################################################################
# DATA_PHENO
################################################################################
# Time, Date and Timestamp
data_pheno$Date <- as.Date("2021-06-28")

# Name of the platform
data_pheno$Platform <- "UCPH"

# Rename the columns for the template
data_pheno <- rename(data_pheno,
                     Unit.ID = Fixture_number,
                     DW_shoot_g = `DW_shoot(g)`,
                     DW_root_g = `DW_root(g)`
                     )

################################################################################
# DATA_IMAGING 
################################################################################
# Time, Date and Timestamp
data_imaging$Timestamp <- data_imaging$`Date and Time`

data_imaging$Date <- sapply(strsplit(as.character(data_imaging$Timestamp), split = " "), '[', 1)
data_imaging$Time <- sapply(strsplit(as.character(data_imaging$Timestamp), split = " "), '[', 2)

data_imaging <- subset(data_imaging, Date >= "2021-06-13") # The plants were put in the platform on the 13rd of june



# Name of the platform
data_imaging$Platform <- "UCPH"

# Rename the columns for the template
data_imaging <- rename(data_imaging,
                       Unit.ID = `Fixture ID:`,
                       Plant_height_mm = `Measurement:`
                           )
```

#### Unit conversions
The data template is only in cm, cm² and g. This step converts the data in the right units.

For the UCPH platform, one variable is in mm.

```{r}
data_imaging$S_Height_cm <- 0.01 * data_imaging$Plant_height_mm
```

# 2. Data template

### A. Data template: plant_info
This dataset contains information about the plant: Unit.ID, genotype, replication, row and column location in the greenhouse, and soil treatment.

```{r Template plant_info, include=FALSE}
plant_info <- data.frame(Unit.ID = factor(design$Unit.ID),
                 Genotype = factor(design$Genotype),
                 Soil = factor(design$Soil),
                 Replication = factor(design$rep),
                 Row = factor(design$nrow),
                 Column = factor(design$ncol),
                 Platform = "UCPH"
                 )
```

### B. Data template: endpoint
This datasets contains information of the end of the experiment (variables at harvest).
It is then linked by the Unit.ID to the plant_info data template.

```{r Template endpoint, include=FALSE}
endpoint <- data.frame(Unit.ID = as.factor(data_pheno$Unit.ID),
                       Time = NA, 
                       Date = date(data_pheno$Date),
                       Timestamp = NA,
                       
                       DW_shoot_g = as.numeric(data_pheno$DW_shoot_g),
                       FW_shoot_g = NA,
                       DW_root_g = as.numeric(data_pheno$DW_root_g),
                       FW_root_g = NA,
                       Leaf_number= NA,
                       Plant_height_cm = NA,
                       DW_plant_g = NA,
                       Root_length_cm = NA,
                       Root_number = NA,
                       Root_angle = NA,
                       Total_wu = NA,
                       DW_seed_g = NA,
                       FW_seed_g = NA,
                       Leaf_area_cmsquared = NA
)
endpoint <- left_join(endpoint, plant_info, by = "Unit.ID")
```

### C. Data template: timeseries
This section in divided in three data templates:

- timeseries

- S_timeseries (variables computed from sideview imaging or image processing)

- T_timeseries (variables computed from topview imaging or image processing)

The time interval between data timestamps varies in each platform.
They are then linked by the Unit.ID to the plant_info data template. 


```{r Template timeseries S_timeseries T_timeseries, include=FALSE}

timeseries <- data.frame(Unit.ID = NA,
                         Time = NA,
                         Date = NA,
                         Timestamp = NA,
                         
                         Manual_Plant_height_cm = NA, 
                         Leaf_number = NA,
                         Wue = NA,
                         Plant_biomass = NA,
                         Ligulated_leaf_number = NA,
                         Plant_emergence = NA,
                         Plant_transpiration = NA,
                         Daily_wu = NA,
                         Soil_water_potential = NA
)
timeseries <- left_join(timeseries, plant_info, by = "Unit.ID")
                      

S_timeseries <- data.frame(Unit.ID = factor(data_imaging$Unit.ID),
                           Timestamp = data_imaging$Timestamp,
                           Date = date(data_imaging$Date),
                           Time = data_imaging$Time,
                           
                           S_Height_cm = as.numeric(data_imaging$S_Height_cm),
                           S_Height_pixel = NA,
                           S_Area_cmsquared = NA,
                           S_Area_pixel = NA,
                           S_Perimeter_cm = NA,
                           S_Perimeter_pixel = NA,
                           S_Convex_hull_area_cmsquared = NA,
                           S_Solidity = NA,
                           S_Compactness = NA,
                           S_Width_cm = NA,
                           S_Width_pixel = NA,
                           S_Leaf_area_cmsquared = NA
)
S_timeseries <- left_join(S_timeseries, plant_info, by = "Unit.ID")


T_timeseries <- data.frame(Unit.ID = NA,
                           Time = NA, 
                           Date = NA,
                           Timestamp = NA,
                           T_Area_cm_squared = NA,
                           T_Area_pixel = NA,
                           T_Perimeter_cm = NA,
                           T_Perimeter_pixel = NA,
                           T_Convex_hull_area_cmsquared = NA,
                           T_Solidity = NA,
                           T_Compactness = NA,
                           T_Roundness = NA,
                           T_Roundness2 = NA,
                           T_Isotropy = NA,
                           T_Eccentricity = NA,
                           T_Rms = NA,
                           T_Sol = NA
)
T_timeseries <- left_join(T_timeseries, plant_info, by = "Unit.ID")

```

### D. NaPPI data templates

- plant_info
- endpoint
- timeseries
- S_timeseries
- T_timeseries

```{r Data template, echo=FALSE}
head(plant_info)
head(endpoint)
head(timeseries)
head(S_timeseries)
head(T_timeseries)
```

# 3. Export the data templates in .txt
Stock the new data sets in a new folder.

```{r Export}
setwd("C:/Users/elise/Documents/Mémoire/Main/Data/Templates/UCPH")

write.table(plant_info, file = "plant_info.txt", sep = "\t", row.names = FALSE, quote = FALSE)
write.table(endpoint, file = "endpoint.txt", sep = "\t", row.names = FALSE, quote = FALSE)
write.table(timeseries, file = "timeseries.txt", sep = "\t", row.names = FALSE, quote = FALSE)
write.table(S_timeseries, file = "S_timeseries.txt", sep = "\t", row.names = FALSE, quote = FALSE)
write.table(T_timeseries, file = "T_timeseries.txt", sep = "\t", row.names = FALSE, quote = FALSE)
```














